---
title: "dplyr"
author: "Dayana Moncada"
date: "8/30/2017"
output: html_document
---

<img src= "LinkedInpic.jpg"/>

International Student from Honduras. First year graduate student at Mercyhurst University in Data Science.
Data Analytics | Political Science & Public Policy Advocacy | Fashion Data Analysis


## Filtering Columns and Rows



When using dplyr, you can filter columns and rows by using select and filter. Let's look at an example using the Lahman baseball database. We first have to load the Lahman and dplyr packages. 

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(dplyr)
```


Now, suppose we would like to see the homerun totals for the 1927 New York Yankees. We could run the following code:

```{r message=FALSE, warning=FALSE}
Batting%>%
  select(playerID,yearID,teamID,HR)%>%
  filter(teamID=="NYA" & yearID==1927)
```

Now, we would like to see the instances where a player had more than 40 HR but less than 60 strikeouts (SO). We could run the following code:

```{r}
Batting%>%
  select(playerID,yearID,teamID,HR,SO)%>%
  filter(HR>=40 & SO<60)
```

To find all instances of Phillies in the 1970s hitting more than 30 homeruns. We could run the following code:
```{r}
Batting%>%
  select(playerID,yearID,teamID,HR)%>%
  filter(HR>30 & teamID== 'PHI' & yearID<=1979 & yearID>=1970)
```

Furthermore, the chart below shows the players hitting more than 50 homeruns arranged from the ones with the most homeruns at the top. We ran this code:
```{r}
Batting%>%
  select(playerID,yearID,teamID,HR)%>%
  filter(HR>50)%>%
  arrange(desc(HR))
```

The chart below shows the instances of a player stricking out less than 10 times. The players have to have at least 400 at-bats (AB). The players with the smallest number of strickouts will be at the top.
```{r}
Batting%>%
  select(playerID,yearID,teamID,SO,AB)%>%
  filter(SO<10 & AB>=400)%>%
  arrange(SO)
```

```{r}
Batting%>%
  select(playerID,yearID,teamID,HR)%>%
  filter(teamID=="NYA" & yearID==1927)
```

```{r}
Batting%>%
  group_by(playerID== 'ruthba01')%>%
  summarize(career_HR=sum(HR))
```



```{r}
Batting%>%
  group_by(playerID)%>%
  summarize(career_HR=sum(HR))%>%
  filter(career_HR>=600)%>%
  arrange(desc(career_HR))
```

```{r}
Batting%>%
  group_by(playerID)%>%
  summarize(HRAvg=mean(HR))%>%
  filter(HRAvg>30)%>%
  arrange(desc(HRAvg))
```


```{r}
Batting%>%
  filter(yearID>=1970)%>%
  group_by(playerID)%>%
  summarize(maxHR=max(HR))%>%
  filter(maxHR>50)%>%
  select(playerID)
```

# ---------- 

